%\chapter*{Preface}
%\addcontentsline{toc}{chapter}{Preface} % Add the preface to the table of contents as a chapter

% TODO: potentialy add personal statement/acknowledgements here



\section*{Course summary}

Energy powers societies all around the world. From mobile phones, to cars, to cook stoves, people use the earth's energy resources, for example crude oil or the energy in wind, to enhance their lifestyles. But these energy systems are also connected to some of the major challenges we are facing at the moment: the climate crisis, air pollution, and providing access to energy for all people all around the world. In order to address these challenges, we need to change the way we source and use energy, so the energy system is environmentally and socially sustainable.\\

In this course, we will learn how we can use computer tools, so-called energy system models, to help us plan a transition to a sustainable energy system. In the first two sessions, we will discuss the energy system and energy system models in general. Afterwards, we will look in more detail at particular techniques used to build and operate these tools. You will also be given a base model and use it yourself to answer questions about how we can generate the electricity we need in future. Based on our discussions and your experience with the model, we will also have a critical look at the strengths and weaknesses of these tools.\\

This course will give you the chance to develop a good understanding of how we use energy, what challenges are related to this, and how to use energy models to address questions about future solutions. It will help you to think more critically about things you hear and read and how to write down your thoughts more precisely.\\

\section*{Course structure}

The course consists of 5 sessions, each including homework that is an integral part of the course. Each session has a separate section in this handbook providing relevant material and tasks. The final coursework will be introduced in session 5 and gives you the opportunity to work independently on your own scientific energy analysis.